|

### 2024-05-19 rmds/manuscript.Rmd
revisions

---
title: |
  Quantifying maternal reproductive output of chondrichthyan fishes
subtitle: |
author:
- name: Alastair V Harry
  affiliation: Fisheries & Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia
  email: alastair.harry@gmail.com
  number: 1
- name: Ivy E Baremore
  affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
  number: 2
- name: Andrew N Piercy
  affiliation: College of Southern Maryland, La Plata MD 20646, United States
  number: 3
toc: false
output:
  word_document: default
  bookdown::pdf_document2:
    pandoc_args:
      - --csl=extras/cjfas.csl
      - --citation-abbreviations=extras/abbreviations.json
      - --template=extras/mytemplate.tex
bibliography: "extras/maternal-output.bib"
keep-latex: true
linestretch: 2 
header-includes:
  - \fontsize{12}{20}
  - \usepackage{lineno}
  - \linenumbers
  - \linenumberfont
  - \renewcommand\linenumberfont{\normalfont\small\sffamily}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{\thepage}
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
editor_options: 
  chunk_output_type: inline
---

@@

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that precision, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could be estimated with greater precision however substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \> 100 estimation of $P_{Max}$ is generally preferable to using a fixed value.

@@

In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015]. Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017]. This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014]. Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stockrecruitment_2013]. This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

@@

Walker [-@walker_reproduction_2005] recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: $$Y_i \sim B(x_i; n_i, P_i)$$ $$E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x)}}$$ where the proportion of individuals in a given length or age category, $x_i$, that are in maternal condition, $Y_i$, is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, $P_i$, multiplied by the total number of individuals in the category, $n_i$. If $x$ is a continuous rather than categorical variable (i.e. $n = 1$), $Y_i$ reduces to a Bernoulli random variable [@zuur_analyzing_2007]. $\alpha$ is the upper asymptote of the curve, controlling the maximum value of $Y$ as $x$ approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when $\alpha = 1$, as is appropriate for most teleost and chondrichthyan populations if modeling maturity [@roa_estimation_1999; @quinn_quantitative_1999]. When $\alpha = 1$, the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows $Y$ to be modeled as a linear function of $x_i$, where $b_0$ and $b_1$ are the intercept and slope, respectively [@zuur_analyzing_2007]. Alternatively, the model can be expressed in more biologically relevant terms as: $$E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} $$ where $x_{50}$ and $x_{95}$ are the sizes or ages at which 50% and 95% of the maximum proportion of individuals ($P_{Max}$) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

@@

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles. Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

@@

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions [@cortes_population_2012]. It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g. annual, biennial), despite observations of plasticity in this trait within discrete populations [@higgs_reproductive_2020]. Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing $P_{Max}$ in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive cycles. To date only a single study on spiny dogfish, *Squalus acanthias*, has statistically estimated $P_{Max}$ from data [@colonello_reproductive_2016], and the feasibility of doing so for a wider range of species has not been investigated.

@@

To assess the feasibility of estimating maternity parameters a simulation study was carried out using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1). Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs. The performance of three methods in estimating 'true' maternity parameters from simulated data was tested by examining precision, bias, and interval coverage. To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters. Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency. All simulation modelling and data analysis was conducted using the R language [@r_core_team_r:_2018] and the Template Model Builder (TMB) R package [@kristensen_tmb:_2016].

@@

Data were generated for the gummy shark, *Mustelus antarcticus*, and school shark, *Galeorhinus galeus*, using parameters from studies conducted on southern Australian populations (Table 1) [@grant_estimation_1979; @kirkwood_gill_1986; @moulton_age_1992; @walker_fishery_1992; @punt_stock_1998; @walker_reproduction_2005; @walker_spatial_2007]. Both species are well-studied and were chosen to be representative of r- and K-selected chondrichthyan life history styles, respectively [@stevens_variable_1999].

@@

Next, maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This process was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.

@@

Two approaches to estimating maternity parameters were compared: the 3PLF itself (3PLF-estimated) and the 3PLF function with a fixed asymptote (3PLF-fixed). Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF maturity). The 3PLF-estimated method was undertaken to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016]. The 3PLF-fixed is the method described by Walker [-@walker_reproduction_2005] and most commonly used in practice. The 2PLF-maternity method was used to examine what effect simply ignoring the upper asmyptote had (i.e. $P_{Max}$ fixed at 1). The 2PLF-maturity method is commonly used to approximate a maternity curve, and tacitly assumed to be similar. The resulting maturity curve can then be weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially [@sedar_sedar_2023]).

@@

Model performance was evaluated in each iteration where 1) the fitted model successfully converged as indicated by a positive definite Hessian matrix, and 2) the estimated parameters $L_{50}'$ and $L_{95}'$ remained within specified bounds ($0 - 10^4$). An iteration was also not attempted if there were no maternal females in the randomly generated sample. Model performance was evaluated by quantifying precision, bias, and interval coverage for $L'_{50}$, $P_{Max}$ and $R_0$. Bias was quantified using the mean relative error, $\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]$ and precision using the mean absolute relative error, $\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|$. Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50% confidence intervals [@rudd_accounting_2018]. For a well-performing model, approximately 50% of parameters would be expected to fall withing the nominal 50% confidence intervals.

@@

An empirical case study was also undertaken to show how maternity functions can be used in practice. Data from two studies of the reproductive biology of the sandbar shark, *Carcharhinus plumbeus*, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods. The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{yr}^{-1}$) with a protracted reproductive cycle lasting longer than a year [@springer_natural_1960]. The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s [@sminkey_age_1995] and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes [@baremore_reproduction_2012; @piercy_histological_2016]. In their study, Baremore and Hale [-@baremore_reproduction_2012] estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle. Piercy *et al* [-@piercy_histological_2016] did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles. To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with $P_{Max}$ estimated and fixed at 0.5 and 0.33, respectively.

@@

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes $\leq$ 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved precision with the 3PLF methods, although at larger sample sizes better precision was achieved by estimating $P_{Max}$ (Figure 3). Overall, $P_{Max}$ was comparatively more challenging to estimate accurately than $L'_{50}$ (Figure S4, Figure S5).

@@

Overall, the 2PLF-maternity method performed the poorest. Ignoring $P_{Max}$ introduced a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3). This manifested in an overestimation of $L'_{50}$ that was exacerbated by gear selectivity effects (Figure S2). In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance. $L_{50}$ could usually be estimated with a higher precision than $L'_{50}$. For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2%, exceeding the performance of the 3PLF-methods at most sample sizes.

The effect of the different methods in ultimately calculating $R_0$ varied considerably across the various scenarios. For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of $R_0$ by 3.4% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1%. (Table S1, Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10). For both species, better precision in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$.

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, $P_{Max}$. Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that precision, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated. Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

Using simulated data to compare the relative performance of the 3PLF-estimated and 3PLF-fixed methods subject to a range of variables showed that it was feasible to estimate $P_{Max}$ from data, but also identified situations where it may be preferable to fix $P_{Max}$. In most scenarios sample sizes of at least 100 were needed for the 3PLF-estimated method to approach or exceed the performance of the 3PLF-fixed method. At sample sizes below this it is therefore advisable to fix $P_{Max}$. Larger sample sizes are desirable before attempting to estimate $P_{Max}$ in species with a triennial or longer reproductive cycle, due to the decline in the performance of the 3PLF-estimated method as $P_{Max}$ decreases.

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of $\geq500$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999]. In the case of the gummy shark this imprecision led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$. Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a sample size of \> 1000. A similar sample size was used by Colonello *et al* [-@colonello_reproductive_2016] to successfully estimate $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].

>>>>

---
title: |
  Quantifying maternal reproductive output of chondrichthyan fishes
subtitle: |
author:
- name: Alastair V Harry
  affiliation: Fisheries & Agriculture Resource Management, Department of Primary Industries and Regional Development, 39 Northside Drive, Hillarys, 6025, WA, Australia
  email: alastair.harry@gmail.com
  number: 1
- name: Ivy E Baremore
  affiliation: MarAlliance, Roatán, Islas de la Bahía, Honduras
  number: 2
- name: Andrew N Piercy
  affiliation: College of Southern Maryland, La Plata MD 20646, United States
  number: 3
toc: false
output:
  bookdown::pdf_document2:
    pandoc_args:
      - --csl=extras/cjfas.csl
      - --citation-abbreviations=extras/abbreviations.json
      - --template=extras/mytemplate.tex
bibliography: "extras/maternal-output.bib"
keep-latex: true
linestretch: 2 
header-includes:
  - \fontsize{12}{20}
  - \usepackage{lineno}
  - \linenumbers
  - \linenumberfont
  - \renewcommand\linenumberfont{\normalfont\small\sffamily}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyfoot[L]{\thepage}
  - \renewcommand{\headrulewidth}{0pt}
  - \renewcommand{\footrulewidth}{0pt}
editor_options: 
  chunk_output_type: inline
---

@@

For the live-bearing and egg-laying class of chondrichthyan fishes a three parameter logistic 'maternity' function with a variable upper asymptote, $P_{Max}$, can be used to predict the average probability of a female giving birth or laying eggs in a season. Although fundamental to calculating the reproductive capacity of a population, few studies report maternity functions, with maturity functions often used as a proxy. Applying logistic models to simulated and empirical data showed that it was feasible to estimate $P_{Max}$ from maternal data and that accuracy, bias, and confidence interval coverage often improved compared to when a fixed value was used. Maturity parameters could also be estimated with greater accuracy; however, substituting them for maternity parameters overestimated lifetime reproductive output. Greater use of maternity functions has the potential to improve calculation of reproductive output in quantitative populations models. In addition to improvements in parameter estimation, this method involves fewer assumptions and enables statistical inferences to be made on frequency of reproduction. This study suggests that for sample sizes \> 100, estimation of $P_{Max}$ is generally preferable to using a fixed value.

@@

In chondrichthyan fishes (sharks, rays, and chimaeras), the process of maturation can be modelled in a similar manner to teleost fishes with a simple 2PLF; however, for the intents and purposes of calculating reproductive capacity, chondrichthyans differ in several ways to most teleosts. Firstly, in many chondrichthyans the duration of the reproductive cycle is longer than a year, meaning the proportion of mature females that will give birth in the population in a given season is less than one [@frisk_life_2005; @dulvy_you_2008; @rigby_patterns_2015]. Secondly, there may be a lag between when females reach sexual maturity and when they begin reproducing [@harry_age_2013; @fujinami_reproductive_2017]. This is confounded by the adopted definition of maturity, which can lead to variable estimates of size and age at maturity [@braccini_determining_2006; @walker_spatial_2007; @montealegre-quijano_sexual_2014]. Finally, because female fecundity is low (usually $\ll$ 100 embryos), total pup or egg production and recruitment are likely to be closely related to the number of reproductively active females, exhibiting little interannual variability [@taylor_stockrecruitment_2013]. This contrasts the often highly variable relationship between stock size and recruitment in teleost fish, and warrants a precise definition and calculation of reproductive output. For these reasons, unlike in teleosts, a maturity function may not be the most appropriate tool to quantify reproductive capacity.

@@

Walker [-@walker_reproduction_2005] recognised the need for a maternity function, as distinct from a maturity function, to accurately quantify reproductive capacity in chondrichthyans. For this purpose he used a non-linear, three-parameter logistic function (3PLF) of the form: $$Y_i \sim B(x_i; n_i, P_i)$$ $$E[Y_i] = n_iP_i = \frac{\alpha}{1 + e^{-(b_0 + b_1x_i)}}$$ where the proportion of individuals in a given length or age category, $x_i$, that are in maternal condition, $Y_i$, is a binomially distributed random variable, with an expected value equal to the probability of an individual from that category being in maternal condition, $P_i$, multiplied by the total number of individuals in the category, $n_i$. If $x$ is a continuous rather than categorical variable (i.e. $n = 1$), $Y_i$ reduces to a Bernoulli random variable [@zuur_analyzing_2007]. $\alpha$ is the upper asymptote of the curve, controlling the maximum value of $Y$ as $x$ approaches infinity. The lesser-used 3PLF reduces to the ubiquitous 2PLF when $\alpha = 1$, as is appropriate for most teleost and chondrichthyan populations if modeling maturity [@roa_estimation_1999; @quinn_quantitative_1999]. When $\alpha = 1$, the function also belongs to the class of generalised linear models; logistic transformation of the dependent variable allows $Y$ to be modeled as a linear function of $x_i$, where $b_0$ and $b_1$ are the intercept and slope, respectively [@zuur_analyzing_2007]. Alternatively, the model can be expressed in more biologically relevant terms as: $$E[Y_i] = P_{Max} \left ( 1 + e^{-ln(19)\left ( \frac{x_i - x_{50}}{x_{95} - x_{50}} \right )}  \right )^{-1} $$ where $x_{50}$ and $x_{95}$ are the sizes or ages at which 50% and 95% of the maximum proportion of individuals ($P_{Max}$) are in maternal condition. Instead of being fit to data on female maturity condition the model is fit to dichotomous data on female maternal condition (non maternal / maternal).

@@

Despite an increase in use, only a small fraction of reproductive studies in recent years have reported maternity functions. Classification of maternal condition is more data-intensive than maturity condition, ideally requiring monthly sampling over a year or longer in order to establish the timing and duration of the female ovarian and uterine cycles, particularly if a there is a resting period between pregnancies. Such data can be difficult and costly to collect for sharks and rays, which are often data-poor and sampled opportunistically. The teleost-oriented foundations of fisheries science have also contributed to the general lack of awareness of maternity functions; Walker (2005) is the sole description on this type of analysis for chondrichthyans and there are no primary literature sources that describe specific methods for chondrichthyan fishes.

@@

Wider use of maternity functions in studies of chondrichthyan reproductive biology is needed to improve calculation of reproductive capacity in the quantitative population assessments that are increasingly being used to support fisheries management and prioritize conservation actions [@cortes_population_2012]. It may also help address the lack of empirical data on the periodicity of chondrichthyan reproductive cycles. Historically, reproductive periodicity has only been determined qualitatively, and has been assumed to be fixed (e.g. annual, biennial), despite observations of plasticity in this trait within discrete populations [@higgs_reproductive_2020]. Information on reproductive periodicity is contained within maternal data, however the existing approach of fixing $P_{Max}$ in maternity functions prevents statistical inference from being made on these data. Changes to the way in which maternity functions are implemented may also contribute to quantifying uncertainty in reproductive frequency and help understand the temporal stability of reproductive periodicity. To date only a single study on spiny dogfish, *Squalus acanthias*, has statistically estimated $P_{Max}$ from data [@colonello_reproductive_2016], and the feasibility of doing so for a wider range of species has not been investigated.

@@

To assess the feasibility of estimating maternity parameters a simulation study was carried out using the 3PLF to generate data for populations with varying reproductive frequencies (Figure 1). Simulations were conducted for a range of gear selectivities and sample sizes reflective of those typically available in reproductive studies and fishery sampling programs. The performance of three methods in estimating 'true' maternity parameters from simulated data was tested by examining accuracy, bias, and interval coverage. To illustrate the effects of misspecifying maternity parameters, lifetime reproductive output was calculated for each simulated population using parameters derived from each of the methods, and by substituting maturity parameters for maternity parameters. Finally, empirical data from previously published studies were reanalysed to illustrate possible strategies for estimating maternity parameters and making inferences on reproductive frequency. All simulation modelling and data analysis was conducted using the R language [Version 4.2.0] [@r_core_team_r:_2022] and the Template Model Builder (TMB) R package [Version 1.9.1] [@kristensen_tmb:_2016].

@@

Data were generated for the gummy shark, *Mustelus antarcticus*, and school shark, *Galeorhinus galeus*, using parameters from studies conducted on southern Australian populations (Table 1) [@grant_estimation_1979; @kirkwood_gill_1986; @moulton_age_1992; @walker_fishery_1992; @punt_stock_1998; @walker_reproduction_2005; @walker_spatial_2007]. Both species are well-studied and were chosen to be representative of fast and slow chondrichthyan life history styles, respectively [@stevens_variable_1999].

@@

Next, dichotomous maturity-at-length data, $\psi(l)$, were randomly generated by simulating from a Bernoulli distribution, $\Psi(l) \sim B(P)$, where $P$ was given by the 2PLF: $$P =  \left ( 1 + e^{-ln(19)\left ( \frac{l - L_{50}}{L_{95} - L_{50}} \right )}  \right )^{-1}$$ This process was then repeated to generate maternity-at-length data, $\Psi'(l)$, using the 3PLF: $$\Psi' \sim B(P)$$ $$P =  P_{Max}\left ( 1 + e^{-ln(19)\left ( \frac{l - L'_{50}}{L'_{95} - L'_{50}} \right )}  \right )^{-1}$$ Finally, 300 iterations were run for each of the 144 unique variables combinations.

@@

Two approaches to estimating maternity parameters were compared: the 3PLF itself (3PLF-estimated) and the 3PLF function with a fixed asymptote (3PLF-fixed). Additionally, two approaches using a 2PLF were also examined, using maternity-at-length data (2PLF-maternity) and maturity-at-length data (2PLF-maturity). The 3PLF-estimated method was undertaken to validate the utility of this model, which has so far been used in only a single study [@colonello_reproductive_2016]. The 3PLF-fixed is the method described by Walker [-@walker_reproduction_2005] and most commonly used in practice. The 2PLF-maternity method was used to examine what effect simply ignoring the upper asmyptote had (i.e. $P_{Max}$ fixed at 1). The 2PLF-maturity method is commonly used to approximate a maternity curve, and tacitly assumed to be similar. The resulting maturity curve can then be weighted by the proportion of gravid females to calculate annual reproductive output (although in practice annual fecundity is more often modified, for example halved for a species suspected to reproduce biennially [@sedar_sedar_2023]).

@@

Model performance was evaluated in each iteration where 1) the fitted model successfully converged as indicated by a positive definite Hessian matrix, and 2) the estimated parameters $L_{50}'$ and $L_{95}'$ remained within specified bounds ($0 - 10^4$). An iteration was also not attempted if there were no maternal females in the randomly generated sample. Model performance was evaluated by quantifying accuracy, bias, and interval coverage for $L'_{50}$, $P_{Max}$ and $R_0$. Bias was quantified using the mean relative error, $\text{MRE} = \text{mean}[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]$ and accuracy using the mean absolute relative error, $\text{MARE} = \text{mean}|[(x_{\text{Estimated}} - x_{\text{True}})/x_{\text{True}}]|$. Interval coverage measures the ability of the model to capture uncertainty and was calculated by taking the proportion of the estimated parameters for each simulation that fell within the 50% confidence intervals [@rudd_accounting_2018]. For a well-performing model, approximately 50% of parameters would be expected to fall withing the nominal 50% confidence intervals.

@@

An empirical case study was also undertaken to show how maternity functions can be used in practice. Data from two studies of the reproductive biology of the sandbar shark, *Carcharhinus plumbeus*, in the Gulf of Mexico and western North Atlantic Ocean were combined and reanalysed using both 3PLF methods. The sandbar shark is a relatively long-lived ($A_{Max}$ = 33 years) and slow growing species ($K = 0.12 \text{ yr}^{-1}$) with a protracted reproductive cycle lasting longer than a year [@springer_natural_1960]. The Gulf of Mexico and western North Atlantic Ocean population has been considered overfished since the late 1970s [@sminkey_age_1995] and during the 2000s detailed studies of sandbar shark reproductive biology and growth were undertaken for stock assessment purposes [@baremore_reproduction_2012; @piercy_histological_2016]. In their study, Baremore and Hale [-@baremore_reproduction_2012] estimated maternity parameters using the 2PLF-maternity method, weighting the final curve by 0.37 based on the proportion of pregnant females observed, most closely aligning with a triennial reproductive cycle. Piercy et al. [-@piercy_histological_2016] did not estimate maternity parameters but also suggested that the average reproductive cycle was likely to be longer than two years based on examination of ovarian follicles. To examine relative support for either a biennial or triennial reproductive cycle, 3PLF-fixed models were fit with $P_{Max}$ estimated and fixed at 0.5 and 0.33, respectively.

@@

For the 3PLF methods, parameter estimates tended to be biased high at lower sample sizes (Figure S2, Figure S3). In some scenarios, particularly with sample sizes $\leq$ 250, fixing the asymptote was effective at reducing bias, although for larger sample sizes neither method was clearly preferable (Figure S2, Figure S3). The accuracy of parameter estimates varied considerably across different combinations of variables used in the simulation and among parameters (Figure S4, Figure S5). Again, in some scenarios with smaller sample sizes $\leq$ 250, fixing the asymptote improved accuracy with the 3PLF methods, although at larger sample sizes better accuracy was achieved by estimating $P_{Max}$ (Figure 3). Overall, $P_{Max}$ was comparatively more challenging to estimate accurately than $L'_{50}$ (Figure S4, Figure S5).

@@

Overall, the 2PLF-maternity method performed the poorest. Ignoring $P_{Max}$ introduced a fixed bias in this parameter that increased in magnitude as reproductive periodicity increased (Figure S3). This manifested in an overestimation of $L'_{50}$ that was exacerbated by gear selectivity effects (Figure S2). In contrast, using a maturity function to approximate the maternity function (2PLF-maturity method) resulted in relatively good performance. $L_{50}$ could usually be estimated with a higher accuracy than $L'_{50}$. For the gummy shark, where $L_{50}$ and $L'_{50}$ were relatively similar, using the 2PLF-maturity method led to a constant underestimate of $L'_{50}$ of approximately 2%, exceeding the performance of the 3PLF-methods at most sample sizes.

The effect of the different methods in ultimately calculating $R_0$ varied considerably across the various scenarios. For school sharks, estimating $P_{Max}$, was the most effective way to minimize bias in most scenarios (Figure 5, Figure S10). Substituting maturity parameters (2PLF-maturity method) led to a median overestimation of $R_0$ by 3.4% across all scenarios, while the 2PLF-maternity method led to a median underestimate of 11.1%. (Table S1, Figure S10). In contrast, for the gummy shark substituting maturity parameters led to a median overestimation of $R_0$ of just 1% across all scenarios which outperformed the 3PLF methods in several instances (Figure 5, Figure S10). Using the 2PLF method for gummy sharks led to a median overestimate of $R_0$ of 22% (Figure S10). For both species, better accuracy in calculating $R_0$ was achieved with the 3PLF methods, with the 3PLF-estimated method performing best in most scenarios tested (Figure 6, Figure S11). For gummy sharks the 3PLF-fixed method performed best at lower sample sizes and lower values of $P_{Max}$.

@@

Relatively few reproductive biology studies have used maternity functions to model maternal reproductive output in chondrichthyan fishes. Where they have been employed the approach has typically been to use a three parameter logistic function with a fixed, user-defined value for the upper asymptote, $P_{Max}$. Here we show that it is feasible to estimate $P_{Max}$ from maternal data, in turn enabling statistical inferences on reproductive periodicity. Applying 3PLF models with estimated and fixed values of $P_{Max}$ to simulated data showed that accuracy, bias, and confidence interval coverage often improved when $P_{Max}$ was estimated. Using a fixed value for $P_{Max}$ in some cases resulted in lower bias at low sample sizes. This study also demonstrated that recruitment can be overestimated when maturity data were used to approximate maternal data and inaccurate if $P_{Max}$ was ignored when estimating maternity parameters. Based on these findings we outline considerations for practitioners using these methods and illustrate how they can provide novel insights into reproductive biology. We conclude by discussing the advantages of adopting this approach to quantifying maternal reproductive output and future directions.

@@

While the estimation of maternity parameters in most simulations was possible, it was difficult to do so accurately. Outcomes of the simulation study suggested that samples sizes of $\geq500$ would be needed to estimate $P_{Max}$ with \<10% MARE for a triennially reproducing species. This is a much larger error than is achievable from conventional logistic maturity analysis [@roa_estimation_1999]. In the case of the gummy shark this inaccuracy led to the 2PLF-maturity method performing comparably or better than the 3PLF methods in ultimately quantifying $R_0$. Relatively wide confidence intervals were also obtained for $P_{Max}$ in the empirical analysis for sandbar sharks, even with a sample size of \> 1000. A similar sample size was used by Colonello *et al*. [-@colonello_reproductive_2016] to estimate plausible values of $P_{Max}$ and maternity parameters for south Atlantic spiny dogfish. These results suggest that data requirements of the 3PLF analyses may be prohibitively large for many chondrichthyans and therefore best suited to use on commercially captured species where large sample sizes can be obtained [@oddone_size_2010; @tribuzio_life_2012].

In light of the data requirements suggested by this study, placing a penalty function on $P_{Max}$ or using a the use of a prior distribution

For some species, and data on pregnant females are particularly sparse, 
the current practice of using maturity parameters as a proxy for maternity parameters will likely still be the only option for numerous data-poor chondrichthyans. From this perspective, the outperformance of the 3PLF-methods by the 2PLF-methods for the gummy shark in several simulations is encouraging. However, the extent to which maturity parameters can provide a good approximation of maternity parameters may be species-specific, depending on how close $L_{50}$ is to $L'_{50}$. While several studies have shown these parameters to be similar [@walker_spatial_2007; @soto-lopez_reproductive_2018], $L'_{50}$ is more frequently shifted to the right of $L_{50}$ [@braccini_determining_2006; @palacios-hernandez_temporal_2020; @montealegre-quijano_sexual_2014; @colonello_reproductive_2016], and this length difference can equate to one or more years. For example, Harry et al. [-@harry_age_2013] found both spot-tail shark, *C. sorrah*, and Australian blacktip shark, *C. tilstoni*, began reproducing the year after reaching sexual maturity. Similarly, Fujinami et al. [-@fujinami_reproductive_2017] found $A'_{50}$ to be 1.4 years older than $A_{50}$ in blue sharks.
